Topic and tone of political news articles in German online media.

### 2018-04-17 paper/writing/structure.tex
change intro

Pluralism of media is an essential principle of democratic societies. For the process of forming opinions, information conveyed by the media, in particular the mass media of journalism, plays a decisive role. They create the preconditions for making the social and cultural diversity of modern societies visible and manageable. For this reason, securing the diversity of opinion-forming processes and preventing concentration processes in the media sector is an essential prerequisite for a functioning democracy. An important question is therefore whether the political reporting of different content providers distinguishes itself and whether this reporting has an influence on the opinion-forming process of the voters. The aim of this paper is to investigate these questions with the use of up-to-date text mining methods. The analysis is based on online news articles from leading German news websites before, during and after the German parliamentary elections in September 2017. The internet as a source of information is gaining in importance, especially in a political context. According to a study on the media coverage of the German-speaking population, nearly 40\% of respondents used the internet at least once a week in 2017 to inform themselves about the current news compared to 34\% in 2016 \citep{vuma_arbeitsgemeinschaft_verbrauchs-_und_medienanalyse_verbrauchs-_2017}.

The new developments in digital media provide new opportunities and challenges for media outlets. The costs of providing and disseminating information have been reduced and the speed of information exchange has increased enormously. At first glance, these developments seem to have a positive effect on the provision of information. However, this also means that the incentive for media to produce primary information declines since the Internet has reduced the marginal utility of content. A study of French news websites from 2013 shows that only 38 percent of the online articles contained original content, with the remainder being copies of those originals \citep{cage_information_2017}. A multitude of factors influence the diversity of online media, including the market structure and the increasing digitization of the industry. Media outlets (TV, Radio, Newspaper) usually operate in two-sided markets, where they serve two customer groups (content consumer and advertiser) that are connected via indirect network effects. The two-sided market structure of the private news market results in news platforms striving to choose their content in such a way that its reach is as large as possible in order to maximize profits from advertising revenues. \citet{steiner_program_1952} concluded, that profit-maximizing media owners may choose to offer the same content, i.e. content aligned with the tastes of the majority. \citet{gabszewicz_press_2001} study the problem of diversity of the political content of newspapers. They find that the maximum differentiation only prevails if the readers sufficiently value the political differentiation between the newspapers the advertising market is small enough. On the other hand, advertising may also have a positive impact on the media, as it enables publishers to report independently of political parties. \citet{ellman_what_2009} analyze a theoretical framework of a two-sided market for newspaper, where readers value accuracy and advertisers value advert-receptive readers. They found that advertising increases the intensity of competition for readers and therefore raises accuracy of media coverage, whereas in the monopolistic case, newspapers under-report news that sufficiently reduces advertiser profits.

Recognizing the importance of media pluralism, the European Commission has endorsed several initiatives to discuss and promote media diversity in the European Union\footnote{https://ec.europa.eu/digital-single-market/en/policies/media-freedom-and-pluralism}. One way to ensure the independence and diversity of the media landscape is through public funded media. In Germany public broadcasting originated in the post-war period and is financed by compulsory fees. To take into account the distinct nature of digital media, the Interstate Broadcasting Agreement (Rundfunkstaatsvertrag) also regulates the scope for action of online services offered by public service broadcasting since 2007. Accordingly, public media are not allowed to distribute purchased content and must - depending on the category of content - set a time limit on its accessibility. In addition, there is a strict advertising ban and prohibition of regional reporting.

With the purpose of examining the media pluralism in the market for German-language online news, this paper analyzes German online news articles about domestic politics. German federal elections took place on 24th of September 2017 and the formation of the government has taken up a period of about five month. The articles considered here dated from 01.06.2017 to 01.03.2018 and thus inform their readers both about the election promises of the parties (before the election) and about the coalition talks (after the election). They therefore make an important contribution to the public's opinion-forming process. The empirical strategy follows a novel approach combining "the two Ts": Topic and Tone \citep{hansen_shocking_2016}. This means that the topics discussed in the articles are identified first (topic), followed by an analysis of how they are discussed with the various news websites (tone). The final step is to check whether the tonality of reporting is reflected in the survey results on voting preferences calculating the cross correlation coefficients. More precisely, the research strategy is as follows:

>>>>

The discussion about the influence of digital media on the political opinion-forming process has gained momentum since the presidential elections in the USA in October 2016 at the latest. Overall, the importance of the Internet as a source of information for political topics has been very strong in recent years. According to a study on the media coverage of the German-speaking population, nearly 40\% of respondents used the internet at least once a week in 2017 to inform themselves about the current news compared to 34\% in 2016 \citep{vuma_arbeitsgemeinschaft_verbrauchs-_und_medienanalyse_verbrauchs-_2017}.

The pluralism nor only of online media, but of media in general is an essential principle of democratic societies. For the process of forming opinions, information conveyed by the media, in particular the mass media of journalism, plays a decisive role. An important question is therefore whether there are convergence tendencies within the mass media - that is, whether popular media tend to report on certain political topics in the same or similar ways and whether this reporting has an influence on the opinion-forming process of the voters. With the purpose of examining the media pluralism in the market for German-language online news, this paper analyzes German online news articles about domestic politics. German federal elections took place on 24th of September 2017 and the formation of the government has taken up a period of about five month. The articles considered here dated from 01.06.2017 to 01.03.2018 and thus inform their readers both about the election promises of the parties (before the election) and about the coalition talks (after the election). They therefore make an important contribution to the public's opinion-forming process. The empirical strategy follows a novel approach combining "the two Ts": Topic and Tone \citep{hansen_shocking_2016}. This means that the topics discussed in the articles are identified first (topic), followed by an analysis of how they are discussed with the various news websites (tone). The final step is to check whether the tonality of reporting is reflected in the survey results on voting preferences calculating the cross correlation coefficients. More precisely, the research strategy is as follows:
